---
title: "Quarto excercise -- BIOS259"
format: html
editor: visual
author:
- name: Aziz Khan
  affiliation: Stanford University, CA, USA
- name: Your Name
  affiliation: Your University Name
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "This is hands-on excrcise for BIOS 259: The Art of Reproducible Science
  – a Stanford Biosciences mini-course on computational reproducibility\n"
tags:
- reproducibility
- notebook
- iris
---

## Introduction

In this **Quarto document**, we'll explore some built-in datasets in R base and create visualizations to analyze the data. The goal is to demonstrate how Quarto combines code and narrative text to produce *reproducible research*.

## Load Data

We'll start by loading the `iris` dataset, which contains measurements of iris flowers.

```{r}
# Load the iris dataset
data(iris)
head(iris)
```

The iris dataset contains measurements of sepal length, sepal width, petal length, and petal width for **`r length(unique(iris$Species))` species** of iris flowers: setosa, versicolor, and virginica.

## Summary Statistics

Let's explore the structure of the iris dataset and summary statistics for each variable.

```{r}
# Explore dataset structure
str(iris)

# Summary statistics
summary(iris)

```

## Data Visualization

### Scatter Plot

We'll create a scatter plot to visualize the relationship between sepal length and sepal width for each species of iris flowers.

```{r, fig.width=8, fig.height=5}
# Scatter plot of sepal length vs. sepal width
library(ggplot2)
iris_scatter_plot <- ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width, color = Species)) +
  geom_point() +
  labs(title = "Scatter Plot",
       x = "Sepal Length", y = "Sepal Width")

# print the plot
iris_scatter_plot
```

### Make the plot publication ready

```{r iris-figure1a, fig.width=8, fig.height=5}
# Load the cowplot
require(cowplot)
# Use the theme from cowplot
iris_scatter_plot <- iris_scatter_plot + theme_cowplot(12)
iris_scatter_plot
```

### Boxplot

Next, we'll create a boxplot to compare the distribution of petal lengths for each species of iris flowers.

```{r iris-figure1b, fig.width=4, fig.height=5}
# Boxplot of petal length by species
iris_box_plot <- ggplot(iris, aes(x = Species, y = Petal.Length, fill = Species)) +
  geom_boxplot() +
  labs(title = "Boxplot",
       x = "Species", y = "Petal Length") + theme_cowplot(12)

iris_box_plot
```

### Density Plot

Finally, let's add a density plot to visualize the distribution of sepal lengths for each species of iris flowers.

```{r iris-figure1c}
# Density plot of sepal length by species
iris_density_plot <- ggplot(iris, aes(x = Sepal.Length, fill = Species)) +
  geom_density(alpha = 0.5) +
  labs(title = "Density Plot",
       x = "Sepal Length", y = "Density") + theme_cowplot(12)
iris_density_plot
```

## Conclusion

In this R Markdown document, we explored the iris dataset available in R base and created visualizations to analyze the data. By combining code and narrative text in R Markdown, we produced a reproducible analysis that can be easily shared and reproduced by others.

The `cowplot` package provides the function `plot_grid()` to arrange plots into a grid and label them.

```{r iris-figure1, fig.width=14, fig.height=4}
#Arranging plots into a grid
plot_grid(iris_scatter_plot, iris_box_plot, iris_density_plot, labels = c('A', 'B','C'),
          ncol=3, rel_widths = c(2,1.5,2))

```

Feel free to modify the code, explore other datasets, or add additional visualizations to further analyze the data!

> **Note:** To generate publication ready figures you can try [ggpubr](https://rpkgs.datanovia.com/ggpubr/index.html)

## R session

A good practice is to print R session to record the versions of the packages used.

```{r}
devtools::session_info()
```
